---
title: "Swiftkey Data Exploration"
author: "BB"
date: '2022-04-21'
output: html_document
---

```{r setup, include=FALSE}
# Libraries
library(stringr)

# Parameters
knitr::opts_chunk$set(echo = TRUE)

raw_data_path <- "../data/en_US.twitter.txt"
zip_data_path <- "../data/Coursera-SwiftKey.zip"
raw_data_url <- "https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip"

# Functions



```

## 1. Download and load data

The data can be downoaded from [Capstone Dataset](https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip). We do optional download to the 'data' folder.

```{r load}

if (!file.exists(raw_data_path)) {
    if (!file.exists(zip_data_path)) {
        download.file(raw_data_url, zip_data_path)
    }
    unzip(zip_data_path,
          exdir = "../data",
          junkpaths = TRUE)
}

con <- file(raw_data_path)

raw_text <- readLines(con)

sample_text <- sample(x = raw_text, 
                      size = floor(0.3 * length(raw_text)))


```


## 2. Length of text lines

For Week 1 quiz we want to find the longest text line:

```{r long_text}

lengths <- str_length(raw_text)

which.max(lengths)
lengths[which.max(lengths)]
raw_text[which.max(lengths)]


```
en_US.twitter.txt: 
- 213 characters
- line number 1.484.357
- Text: It's time for you to give me a little bit of lovin'ï¼ˆã\u0081•ã\u0081\u0081ã\u0081¡ã‚‡ã\u0081£ã\u0081¨ã\u0081¯ã\u0081‚ã\u0081ªã\u0081Ÿã\u0081®æ„›ã‚’ã\u0081¡ã‚‡ã\u0081†ã\u0081 ã\u0081„ï¼‰Baby, hold me tight and do what I tell youï¼\u0081ï¼ˆãƒ™ã‚¤ãƒ“ãƒ¼æŠ±ã\u0081\u008dã\u0081—ã‚\u0081ã\u0081¦ç§\u0081ã\u0081Œè¨€ã\u0081†ã‚ˆã\u0081†ã\u0081«ï¼\u0081ï¼‰

en_US.blogs.txt:
- 483.415 characters
- line number 40.835
- Text: UPDATE AS OF 11:30 A.M. EDT, MONDAY, APRIL 11: No damage to Japan's nuclear power plants was reported today after another strong aftershock hit the ...

en_US.news.txt:
- 14.556 characters
- line number 5.760
Text: Assets Average Average Compound Fund ($Mil) Maturity Yield(%) Yield(%) AZL Money Market Fund 858.9 36 0.00 0.00 Amer Century Capital Presv Fund/In 2,878.9 ...




## 3. "love/hate" Ratio

For week 1 quiz we want in en_US.twitter.txt to 'divide the number of lines where the word "love" (all lowercase) occurs by the number of lines the word "hate" (all lowercase) occurs'

```{r love hate ratio}
love_lines <- str_subset(raw_text, "love")
hate_lines <- str_subset(raw_text, "hate")
    
length(love_lines) / length(hate_lines)    
```

The love/hate ratio in en_US.twitter.txt is 4.108592.
