---
title: "Next Word Modelling (week 2)"
author: "BB"
date: '2022-04-28'
output: html_document
---

```{r setup, include=FALSE}
# Libraries
library(stringr)
library(tidyverse)
library(caret)
library(recipes)


# Parameters
knitr::opts_chunk$set(echo = TRUE)

# Data 
word <- read_rds("../data/en_US.twitter_word.rds.gz")
gram2 <- read_rds("../data/en_US.twitter_gram2.rds.gz")
gram3 <- read_rds("../data/en_US.twitter_gram3.rds.gz")
gram4 <- read_rds("../data/en_US.twitter_gram4.rds.gz")

# Functions

source("./helper_functions.R")


```

## 1. Dataset Building

Datasets needed for a basic n-gram model for predicting the next word based on the previous 1, 2, or 3 words.

```{r datasets}

swiftkey_1 <- as.data.frame(word) %>% 
    group_by(word) %>% 
    summarize(freq = n()) %>% 
    arrange(-freq) %>% 
    filter(word != "")
    

swiftkey_2 <- as.data.frame(gram2) %>% 
    separate(gram2, c("w1", "w2"), " ") %>% 
    group_by(w1,w2) %>% 
    summarize(freq = n()) %>% 
    arrange(-freq)

swiftkey_3 <- as.data.frame(gram3) %>% 
    separate(gram3, c("w1", "w2", "w3"), " ") %>% 
    group_by(w1,w2,w3) %>% 
    summarize(freq = n()) %>% 
    arrange(-freq)

swiftkey_4 <- as.data.frame(gram4) %>% 
    separate(gram4, c("w1", "w2", "w3", "w4"), " ") %>% 
    group_by(w1,w2, w3, w4) %>% 
    summarize(freq = n()) %>% 
    arrange(-freq)

swiftkey <- list("1" = swiftkey_1,
                 "2" = swiftkey_2,
                 "3" = swiftkey_3,
                 "4" = swiftkey_4)


```


## 2. Prediction

Predict next word from one given word, based on bigrams:

```{r bigram prediction}

predict2 <- function(corpus, word1) {

    set <- corpus[corpus$w1 == word1,]
    
    n_cases <- sum(set$freq, na.rm = TRUE)
    
    probs <- set %>% 
        mutate(p = round(freq / n_cases,2)) %>% 
        select(-freq) %>% 
        arrange(-p)
    
    probs$w2[1]
}

predict2(swiftkey_2, "you")
predict2(swiftkey_2, "to")
predict2(swiftkey_2, "me")
predict2(swiftkey_2, "oger")


```

Predict next words form a given two-word phrase based on trigrams:

```{r trigram prediction}

predict3 <- function(corpus, word1, word2) {
    
    set <- corpus %>% 
        filter(w1==word1, w2==word2)
    
    n_cases <- sum(set$freq, na.rm = TRUE)
    
    probs <- set %>% 
        mutate(p = round(freq / n_cases,2)) %>% 
        select(-freq)
    
    probs$w3[1]
}

predict3(swiftkey_3, "don't", "you")
predict3(swiftkey_3, "love", "to")
predict3(swiftkey_3, "make", "me")
predict3(swiftkey_3, "big", "oger")


```

Predict the next words form a given three-word phrase based on tetragrams:

```{r tetragram prediction}

predict4 <- function(corpus, word1, word2, word3) {
    
    set <- corpus %>% 
        filter(w1==word1, w2==word2, w3 == word3)
    
    n_cases <- sum(set$freq, na.rm = TRUE)
    
    probs <- set %>% 
        mutate(p = round(freq / n_cases,2)) %>% 
        select(-freq)
    
    probs$w4[1]

}

predict4(swiftkey_4,"why", "don't", "you")
predict4(swiftkey_4,"i", "love", "to")
predict4(swiftkey_4,"you", "make", "me")
predict4(swiftkey_4,"the", "big", "oger")


```

Predict from a given phrase by combination of above and handle unseen ngrams:

```{r predict phrase}

predict_phrase <- function(corpus, phrase) {
    
    corpus <- swiftkey
    phrase <- "the big oger"
    
    clean_phase <- clean_text(phrase)
    final_words <- last_3_words(clean_phase)
    
    pred <- case_when(
        length(final_words) == 3 ~
            predict4(corpus[["4"]], 
                     final_words[1], 
                     final_words[2], 
                     final_words[3]),
        length(final_words) == 2 ~
            predict3(corpus[["3"]], 
                     final_words[1], 
                     final_words[2]),
        length(final_words) == 1 ~
            predict2(corpus[["2"]], 
                     final_words[1]),
        TRUE ~ NA
    )
    
    pred
    
    ifelse(is.na(pred), corpus[["1"]]$word[1], pred)
}

predict_phrase(swiftkey, "me")
predict_phrase(swiftkey, "make me")
predict_phrase(swiftkey, "you make me")
predict_phrase(swiftkey, "the big oger")

```

